---
title: "Lysimachia ciliata | Range-wide distribution under future climate projections"
author: "Mark Buckner"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Packages
```{r Load pkgs, message=FALSE}
library(spatial)
library(rgdal)
library(spThin)
library(ENMeval, attach.required = T)
library(rmaxent)
library(raster)
library(tidyverse)
library(lubridate)
library(mapdata)

#Needed a multistep process to get rJava to install. 
#First: https://stackoverflow.com/questions/63830621/installing-rjava-on-macos-catalina-10-15-6
#This allowed for rJava to install without an error but crash when running. Alert said Java 6 required.
#Second: https://www.l3harrisgeospatial.com/Support/Self-Help-Tools/Help-Articles/Help-Articles-Detail/ArtMID/10220/ArticleID/23780/macOS-Catalina-1015-ENVIIDL-and-Legacy-Java-6-Dependencies
#Allows for Java 6 install and normal rJava function

library(rJava)
```

#### Functions

```{r Functions}
#Function code: https://babichmorrowc.github.io/post/2019-04-12-sdm-threshold/
sdm_threshold <- function(sdm, occs, type = "mtp", binary = FALSE){
  occPredVals <- raster::extract(sdm, occs)
  if(type == "mtp"){
    thresh <- min(na.omit(occPredVals))
  } else if(type == "p10"){
    if(length(occPredVals) < 10){
      p10 <- floor(length(occPredVals) * 0.9)
    } else {
      p10 <- ceiling(length(occPredVals) * 0.9)
    }
    thresh <- rev(sort(occPredVals))[p10]
  }
  sdm_thresh <- sdm
  sdm_thresh[sdm_thresh < thresh] <- NA
  if(binary){
    sdm_thresh[sdm_thresh >= thresh] <- 1
  }
  return(sdm_thresh)
}

#futurePred() : Predicts distribution over a certain model raster stack and plots output. Written to be used with list.files() and sapply().
#file = file name of raster stack; model = training model; names = envnames of predictor layers, vector; host = (optional) two letter abbr. of host, character; drop = names of layers to drop from final raster stack, vector.
futurePred <- function(file, model, envnames, host, drop) {
  path <- paste("./", file, sep = "")
  m <- str_extract(file, "_.*_")
  m <- str_sub(m, 2, -2)
  
  if(missing(host)) {
    r <- stack(path)
  } else {
    yr <- str_extract(file, "\\d+")
    ssp <- str_extract(file, "ssp.*")
    hpath <- paste("../out/", host, "/", host, "_SDM_ME_rw_mean20", yr, ssp, sep = "")
    r <- stack(path, hpath)
  }
  
  names(r) <- envnames
  
  if(!missing(drop)) {
    r <- dropLayer(r, drop)
  }
  
  pred <- predict(model, r)
  
  plot(pred, main = m)
  map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")
  
  return(pred)
}

#futureSave() : Calculates the mean distribution for all models at a particular time and ssp, saves results of mean and p10 threshold, and plots.
#pred = mean prediction; spp = two letter species name abbr., character; year = YYYY; ssp = CMIP6 scenario
futureSave <- function(pred, spp, year, ssp, thres) {
  pred.s <- stack(pred)
  pred.m <- calc(pred.s, mean)

  plot(pred.m, main = paste(year, ssp, "mean", sep = "_"))
  map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")
  
  name <- paste("../../pred/out/",spp, "/", spp, "_SDM_ME_rw_", "mean", year, ssp, ".tif", sep = "")
  
  writeRaster(pred.m, filename = name, options="INTERLEAVE=BAND", overwrite=TRUE)

  #10th percentile training presence
  if(!missing(thres)) {
    p10 <- pred.m
    p10[p10 < thres] <- NA 
    plot(p10, main = paste(year, ssp, "p10", sep = "_"))
    map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")
  
    name <- paste("../../pred/out/",spp, "/", spp, "_SDM_ME_rw_", "p10_", year, ssp, ".tif", sep = "")
  
    writeRaster(p10, filename = name, options="INTERLEAVE=BAND", overwrite=TRUE)
  }
}
```

## Load Occurance Data and Predictors

Occurrence and predictor data processed in `occ_pred_data_rw.Rmd`. Occurrence data needs to have the unnecessary fields removed and projected to the project projection (`+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs`) to create a `matrix array`. 

```{r occ, message=FALSE}
#Occurrences
occ <- read_csv("../../occ/Lc_RW70.csv")

occs <- select(occ, lat, lon) %>% 
  mutate("spp" = "Lc")

occp <- rgdal::project(as.matrix(occs[,c("lon","lat")]), proj = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
```

The predictors are described below. Only bold predictors are included after accounting for collinearity. All other variables were highly correlated r > 0.7 with one of the other variables and were excluded. Variables were chosen based on the network of correlation and the perceived importance.

Bioclimatic variables (USGS descriptions available [here](https://pubs.usgs.gov/ds/691/ds691.pdf))

 - Bio 1 : Mean Annual Temperature
 - Bio 2 : Annual Mean Diurnal Range
 - Bio 3 : Isothermality 
 - **Bio 4 : Temperature Seasonality**
 - **Bio 5 : Max Temperature of Warmest Month**
 - Bio 7 : Annual Temperature Range
 - **Bio 8 : Mean Temperature of Wettest Quarter**
 - Bio 9 : Mean Temperature of Driest Quarter
 - Bio 10 : Mean Temperature of Warmest Quarter
 - Bio 11 : Mean Temperature of Coldest Quarter
 - **Bio 12 : Annual Precipitation**
 - Bio 13 : Precipitation of Wettest Month
 - Bio 14 : Precipitation of Driest Month
 - **Bio 15 : Precipitation Seasonality**
 - Bio 16 : Precipitation of Wettest Quarter
 - Bio 17 : Precipitation of Driest Quarter
 - **Bio 18 : Precipitation of Warmest Quarter**
 - Bio 19 : Precipitation of Coldest Quarter
 
Topography

 - **DEM : SRTM at 5 arc-minute resolution downloaded from WorldClim**
 
Other

 - Solar Radiation : Mean solar radiation during meteorological summer (June, July, August)
 - **Soil Sand Content : Soil sand content (%) at 0cm of depth**
 - **Soil Clay Content : Soil clay content (%) at 0cm of depth**
 - **Soil pH : Soil pH x 10 at 0cm of depth**

General Circulation Models (GCMs)

 - bcc    : BCC-CSM2-MR | [Beijing Climate Center Climate System Model](https://gmd.copernicus.org/articles/12/1573/2019/)
 - cnrm_c : CNRM-CM6-1 | [CNRM/CERFACS](http://www.umr-cnrm.fr/cmip6/spip.php?article11)
 - cnrm_e : CNRM-ESM2-1 | [CNRM/CERFACS](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2019MS001791)
 - can    : CanESM5 | [Canadian Earth System Model](https://gmd.copernicus.org/articles/12/4823/2019/gmd-12-4823-2019.html)
 - ipsl   : IPSL-CM6A-LR | [Institut Pierre-Simon Laplace](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2019MS002010)
 - miroc  : MIROC-ES2L | [](https://gmd.copernicus.org/articles/13/2197/2020/)
 - miroc6 : MIROC6 | [](https://gmd.copernicus.org/articles/12/2727/2019/)
 - mri    : MRI-ESM2-0 | [Meteorological Research Institute](https://www.jstage.jst.go.jp/article/jmsj/advpub/0/advpub_2019-051/_article/-char/en)
 

```{r pred, message = FALSE}
#Predictors
#Load and crop baseline climate conditions. Using most recent available data.
env <- stack("../../pred/pred_stack_rw/baseline.tif")

#worldClim
#envnames <- c("bio4", "bio8", "bio10", "bio12", "bio15", "bio18", "dem", "sand", "clay", "pH")

envnames <- c("bio4", "bio5", "bio8", "bio12", "bio15", "bio18", "dem", "sand", "clay", "pH")

names(env) <- envnames

plot(env[[1:nlayers(env)]])
```

```{r occ plot}
plot(env[[1]], main="Temp. Seasonality | L. ciliata Occ.")
points(occp, pch = 16, cex = 0.3)
```

### Spatial Thinning

```{r spThin}
thinned <-
  thin( loc.data = occs, 
        lat.col = "lat", long.col = "lon", 
        spec.col = "spp", 
        thin.par = 10, reps = 100, 
        locs.thinned.list.return = TRUE, 
        write.files = TRUE, 
        max.files = 5, 
        out.dir = "../../occ/Lc_thinned_full/", out.base = "Lc_thinned", 
        write.log.file = TRUE,
        log.file = "Lc_thinned_full_log_file.txt" )

plotThin(thinned)
```

```{r load thin, message=FALSE}
#Load thinned dataset in place of original
occs <- read_csv("../../occ/lc_thinned_full/Lc_thinned_thin1.csv") %>% 
  select(lat,lon)
  
occs <- rgdal::project(as.matrix(occs[,c("lon","lat")]), proj = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
```

## MESS

```{r MESS, message = FALSE, warning=FALSE}
occs.z <- raster::extract(env, occs)
occs.sim <- similarity(env, occs.z)
occs.mess <- occs.sim$similarity_min
occs.sp <- sp::SpatialPoints(occs)

myScale <- seq(cellStats(occs.mess, min), cellStats(occs.mess, max), length.out = 100)
rasterVis::levelplot(occs.mess, main = "Environmental similarity", at = myScale, margin = FALSE) + 
  latticeExtra::layer(sp.points(occs.sp, col="black"))
```

### Random Background Points

```{r Bkg points}
(points <- sum(!is.na(getValues(env[[1]])))*0.4)


bg <- dismo::randomPoints(env, n = round(points)) %>% as.data.frame()
colnames(bg) <- colnames(occs)

plot(env[[1]], main="Mean ann. temp. | Background Points")
points(bg, pch = 20, cex = 0.05)
```

### Partition for Model Evaluation

```{r partition, message=FALSE}
block <- get.block(occs, bg)
table(block$occs.grp)

evalplot.grps(pts = occs, pts.grp = block$occs.grp, envs = env) + 
  ggplot2::ggtitle("Block partitions: occurrences")
```

### Env. Similarity of Partions

```{r Env Sim, message=FALSE, warning=FALSE}
occs.z <- cbind(occs, raster::extract(env, occs))
bg.z <- cbind(bg, raster::extract(env, bg))
evalplot.envSim.hist(sim.type = "mess", ref.data = "occs", occs.z = occs.z, bg.z = bg.z, 
                     occs.grp = block$occs.grp, bg.grp = block$bg.grp)

evalplot.envSim.hist(sim.type = "most_diff", ref.data = "occs", occs.z = occs.z, bg.z = bg.z, 
                     occs.grp = block$occs.grp, bg.grp = block$bg.grp)

evalplot.envSim.map(sim.type = "mess", ref.data = "occs", envs = env, occs.z = occs.z, 
                    bg.z = bg.z, occs.grp = block$occs.grp, bg.grp = block$bg.grp, bb.buf = 7)
```

### Tune model

```{r Tune}
e.mx <- ENMevaluate(occs = occs, envs = env, bg = bg, 
                    algorithm = 'maxent.jar', partitions = 'block', parallel = TRUE, 
                    tune.args = list(fc = c("L","LQ","LQH","H"), rm = 1:5))

e.mx
```

### Model turning results and Selection

```{r model selection}
evalplot.stats(e = e.mx, stats = c("AICc"), color = "fc", x.var = "rm", 
               error.bars = FALSE)

evalplot.stats(e = e.mx, stats = c("auc.val"), color = "fc", x.var = "rm", 
               error.bars = FALSE)
```

```{r AICc}
res <- eval.results(e.mx)
kable(res)

opt.aicc <- res %>% 
  filter(delta.AICc <= 2) %>% 
  slice_max(auc.val.avg)
kable(opt.aicc)
```

```{r Resp curves, comment=FALSE, message=FALSE}
par(mar = c(4, 2, 0.5, 1)) 
dismo::response(eval.models(e.mx)[[opt.aicc$tune.args]])
par(mar = c(4, 2, 1, 1)) 
plot(eval.models(e.mx)[[opt.aicc$tune.args]])
par(mar = c(5.1, 4.1, 4.1, 2.1))
```

### Prediction

```{r Prediction}
pred.seq <- eval.predictions(e.mx)[[opt.aicc$tune.args]]
plot(pred.seq)
map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")

writeRaster(pred.seq, "../../pred/out/Lc/Lc_SDM_ME_rw_baseline.tif", overwrite = TRUE)
```

### Threshold

#### Minimum training presence 

```{r MTP}
mtp <- sdm_threshold(pred.seq, occs, type = "mtp")
plot(mtp, xlim = c(-130, -55))
map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")

writeRaster(mtp, "../../pred/out/Lc/Lc_SDM_ME_rw_mtp_baseline.tif", overwrite = TRUE)
```

#### 10th percentile training presence

```{r baseline p10}
p10 <- sdm_threshold(pred.seq, occs, type = "p10")
plot(p10)
map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")

p10.b <- sdm_threshold(pred.seq, occs, type = "p10", binary = TRUE)
plot(p10.b)
map('worldHires', fill=FALSE, add=TRUE, xlim = c(-127.75, -55), ylim = c(30.83333, 62.25), lforce = "e")

writeRaster(p10, "../../pred/out/Lc/Lc_SDM_ME_rw_p10_baseline.tif", overwrite = TRUE)
```

### Future Climate Models

Select optimal model as determined above using `ENMeval`. This model will be used to predict over the GCM predictions for years 2041-2060 and 2081-2100 (referred to as 2050 and 2090 respectfully) as well as each of the four Shared Socioeconomic Pathways (SSP) from CMIP6.

```{r select model} 
for (i in 1:length(e.mx@models)) {
  if(names(e.mx@models[i]) == as.character(opt.aicc[,3])) {
    model <- e.mx@models[[i]]
    print(paste("Opt. model : ", names(e.mx@models[i]), sep = ""))
  }
}

(thres <- min(na.omit(as.vector(p10))))
```

#### 2041 - 2060

##### 2050 | ssp126

```{r 2050ssp126, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "50.*ssp126.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2050 | ssp126 | Mean and Threshold

```{r 2050ssp126 mean}
futureSave(pred, "Lc", "2050", "ssp126", thres = thres)
```

##### 2050 | ssp245

```{r 2050ssp245, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "50.*ssp245.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2050 | ssp245 | Mean and Threshold

```{r 2050ssp245 mean}
futureSave(pred, "Lc", "2050", "ssp245", thres = thres)
```

##### 2050 | ssp370

```{r 2050ssp370, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "50.*ssp370.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2050 | ssp370 | Mean and Threshold

```{r 2050ssp370 mean}
futureSave(pred, "Lc", "2050", "ssp370", thres = thres)
```

##### 2050 | ssp585

```{r 2050ssp585, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "50.*ssp585.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2050 | ssp585 | Mean and Threshold

```{r 2050ssp585 mean}
futureSave(pred, "Lc", "2050", "ssp585", thres = thres)
```



#### 2081 - 2100

##### 2090 | ssp126

```{r 2090ssp126, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "90.*ssp126.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```


```{r, eval=FALSE, include=FALSE}
bbc <- stack("../../Data/Pred/wc5_future/2090_ssp126/bbc_126_90.tif")
names(bbc) <- names(env)
bbc.p <- predict(eval.models(e.mx)[[opt.aicc$tune.args]], bbc)
plot(bbc.p)
map('worldHires', fill=FALSE, add=TRUE)


cnrm_c <- stack("../../Data/Pred/wc5_future/2090_ssp126/cnrm_c_126_90.tif")
names(cnrm_c) <- names(env)
cnrm_c.p <- predict(e.mx@models$fc.H_rm.1, cnrm_c)
plot(cnrm_c.p)
map('worldHires', fill=FALSE, add=TRUE)

cnrm_e <- stack("../../Data/Pred/wc5_future/2090_ssp126/cnrm_e_126_90.tif")
names(cnrm_e) <- names(env)
cnrm_e.p <- predict(e.mx@models$fc.H_rm.1, cnrm_e)
plot(cnrm_e.p)
map('worldHires', fill=FALSE, add=TRUE)

can <- stack("../../Data/Pred/wc5_future/2090_ssp126/can_126_90.tif")
names(can) <- names(env)
can.p <- predict(e.mx@models$fc.H_rm.1, can)
plot(can.p)
map('worldHires', fill=FALSE, add=TRUE)


ipsl <- stack("../../Data/Pred/wc5_future/2090_ssp126/ipsl_126_90.tif")
names(ipsl) <- names(env)
ipsl.p <- predict(e.mx@models$fc.H_rm.1, ipsl)
plot(ipsl.p)
map('worldHires', fill=FALSE, add=TRUE)


miroc <- stack("../../Data/Pred/wc5_future/2090_ssp126/miroc_126_90.tif")
names(miroc) <- names(env)
miroc.p <- predict(e.mx@models$fc.H_rm.1, miroc)
plot(miroc.p)
map('worldHires', fill=FALSE, add=TRUE)


miroc6 <- stack("../../Data/Pred/wc5_future/2090_ssp126/miroc6_126_90.tif")
names(miroc6) <- names(env)
miroc6.p <- predict(e.mx@models$fc.H_rm.1, miroc6)
plot(miroc6.p)
map('worldHires', fill=FALSE, add=TRUE)


mri <- stack("../../Data/Pred/wc5_future/2090_ssp126/mri_126_90.tif")
names(mri) <- names(env)
mri.p <- predict(e.mx@models$fc.H_rm.1, mri)
plot(mri.p)
map('worldHires', fill=FALSE, add=TRUE)
```

##### 2090 | ssp126 | Mean and Threshold

```{r 2090ssp126 mean}
futureSave(pred, "Lc", "2090", "ssp126", thres = thres)
```


```{r, eval=FALSE, include=FALSE}
pred.s <- stack(pred)
pred.m <- calc(pred.s, mean)

plot(pred.m)
map('worldHires', fill=FALSE, add=TRUE)

p10 <- sdm_threshold(pred.m, occs, type = "p10")
plot(p10)
map('worldHires', fill=FALSE, add=TRUE)
```

##### 2090 | ssp245

```{r 2090ssp245, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "90.*ssp245.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2090 | ssp245 | Mean and Threshold

```{r 2090ssp245 mean}
futureSave(pred, "Lc", "2090", "ssp245", thres = thres)
```

##### 2090 | ssp370

```{r 2090ssp370, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "90.*ssp370.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2090 | ssp370 | Mean and Threshold

```{r 2090ssp370 mean}
futureSave(pred, "Lc", "2090", "ssp370", thres = thres)
```

##### 2090 | ssp585

```{r 2090ssp585, warning = FALSE}
setwd("../../pred/pred_stack_rw/")
mdls <- list.files(pattern = "90.*ssp585.tif")

par(mfrow = c(2,4))
pred <- sapply(mdls, futurePred, envnames = envnames, model = model)
par(mfrow = c(1,1))
```

##### 2090 | ssp585 | Mean and Threshold

```{r 2090ssp585 mean}
futureSave(pred, "Lc", "2090", "ssp585", thres = thres)
```
