---
title: "Macropis Nuda SDM | MaxEnt"
author: "Mark Buckner"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load pkgs, message=FALSE}
library(spatial)
library(rgdal)
library(ENMeval, attach.required = T)
library(rmaxent)
library(raster)
library(tidyverse)
library(lubridate)
library(knitr)
```

### Load Occurance Data and Predictors

Occurrence and predictor data processed in `occ_pred_data_1km.Rmd`. Occurrence data needs to have the unique identifiers removed and projected to the project projection (`+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs`) to create a `matrix array`. The predictors are described below.

Bioclimatic variables (USGS descriptions available [here](https://pubs.usgs.gov/ds/691/ds691.pdf))

 - Bio 1 : Mean Annual Temperature
 - Bio 2 : Annual Mean Diurnal Range
 - Bio 3 : Isothermality 
 - Bio 4 : Temperature Seasonality
 - Bio 5 : Max Temperature of Warmest Month
 - Bio 7 : Annual Temperature Range
 - Bio 8 : Mean Temperature of Wettest Quarter
 - Bio 9 : Mean Temperature of Driest Quarter
 - Bio 10 : Mean Temperature of Warmest Quarter
 - Bio 11 : Mean Temperature of Coldest Quarter
 - Bio 12 : Annual Precipitation
 - Bio 13 : Precipitation of Wettest Month
 - Bio 14 : Precipitation of Driest Month
 - Bio 15 : Precipitation Seasonality
 - Bio 16 : Precipitation of Wettest Quarter
 - Bio 17 : Precipitation of Driest Quarter
 - Bio 18 : Precipitation of Warmest Quarter
 - Bio 19 : Precipitation of Coldest Quarter
 
Topographical and soils data

 - DEM : Elevation in m
 - Proximity to Water : Arbitrary 
 - Soil Water Content : 0cm depth
 - Soil Sand Content : 0cm depth
 
Host Plant

 - 1km resolution *Lysimachia ciliata* SDM
 - 1km resolution *Apocynum androsaemifolium* SDM (Commonly visits to nectar)

```{r Occ and pred, message=FALSE}
#Predictors
#env <- stackOpen("../../pred/pred_stack_ne/stack")
env <- stack("../../pred/pred_stack_ne/baseline.tif")
names(env) <- c(paste("bio", seq(1,19), sep = ""), "dem", "wprox", "sand", "h20")

lc <- raster("../../pred/out/Lc/Lc_SDM_ME_1km.tif")
aa <- raster("../../pred/out/Aa/Aa_SDM_ME_1km.tif")

env <- addLayer(env, lc, aa)

plot(env[[1:12]])
plot(env[[13:nlayers(env)]])

#Occurrences
occ <- read_csv("../../occ/Mn_NE70.csv")

occs <- select(occ, lat, lon)
occs <- rgdal::project(as.matrix(occs[,c("lon","lat")]), proj = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

plot(env[[1]], main="Mean ann. temp. | M. nuda Occ.")
points(occs, pch = 16, cex = 0.5)


#Load spThin dataset from Lc_Mn_SDM_1km_RF.Rmd
occs <- read_csv("../../occ/Mn_NE_thinned_full/Mn_ne_thinned_thin1.csv") %>% 
  select(lat,lon)
  
occs <- rgdal::project(as.matrix(occs[,c("lon","lat")]), proj = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

plot(env[[1]], main="Mean ann. temp. | M. nuda Occ.")
points(occs, pch = 16, cex = 0.5)
```

### MESS

```{r MESS, message=FALSE}
occs.z <- raster::extract(env, occs)
occs.sim <- similarity(env, occs.z)
occs.mess <- occs.sim$similarity_min
occs.sp <- sp::SpatialPoints(occs)

myScale <- seq(cellStats(occs.mess, min), cellStats(occs.mess, max), length.out = 100)
rasterVis::levelplot(occs.mess, main = "Environmental similarity", at = myScale, margin = FALSE) + 
  latticeExtra::layer(sp.points(occs.sp, col="black"))
```

### Random Background Points

```{r Bkg points}
(points <- sum(!is.na(getValues(env[[1]])))*0.15)
bg <- dismo::randomPoints(env, n = round(points)) %>% as.data.frame()
colnames(bg) <- colnames(occs)

plot(env[[1]])
points(bg, pch = 20, cex = 0.3)
```

### Partition for Model Evaluation

```{r partition, message=FALSE}
block <- get.block(occs, bg)

evalplot.grps(pts = occs, pts.grp = block$occs.grp, envs = env) + 
  ggplot2::ggtitle("Spatial Block partitions: occurrences")
```

### Env. Similarity of Partions

```{r Env Sim}
occs.z <- cbind(occs, raster::extract(env, occs))
bg.z <- cbind(bg, raster::extract(env, bg))
evalplot.envSim.hist(sim.type = "mess", ref.data = "occs", occs.z = occs.z, bg.z = bg.z, 
                     occs.grp = block$occs.grp, bg.grp = block$bg.grp)

#evalplot.envSim.hist(sim.type = "most_diff", ref.data = "occs", occs.z = occs.z, bg.z = bg.z, 
#                     occs.grp = block$occs.grp, bg.grp = block$bg.grp)

evalplot.envSim.map(sim.type = "mess", ref.data = "occs", envs = env, occs.z = occs.z, 
                    bg.z = bg.z, occs.grp = block$occs.grp, bg.grp = block$bg.grp, bb.buf = 7)
```

### Tune model

```{r Tune}
library(rJava)

e.mx <- ENMevaluate(occs = occs, envs = env, bg = bg, 
                    algorithm = 'maxent.jar', partitions = 'block', parallel = TRUE, 
                    tune.args = list(fc = c("L","LQ","LQH","H"), rm = 1:5))

e.mx
```

### Model turning results and Selection

```{r model selection}
evalplot.stats(e = e.mx, stats = c("AICc"), color = "fc", x.var = "rm", 
               error.bars = FALSE)

evalplot.stats(e = e.mx, stats = c("auc.val"), color = "fc", x.var = "rm", 
               error.bars = FALSE)
```

```{r AICc}
res <- eval.results(e.mx)
kable(res)

opt.aicc <- res %>% 
  filter(delta.AICc <= 2) %>% 
  slice_max(auc.val.avg)
kable(opt.aicc)
```

```{r Resp curves, message=FALSE}
par(mar = c(1, 1, 1, 1))
dismo::response(eval.models(e.mx)[[opt.aicc$tune.args]])
plot(eval.models(e.mx)[[opt.aicc$tune.args]])
```

### Prediction

```{r Prediction, message = FALSE}
pred.seq <- eval.predictions(e.mx)[[opt.aicc$tune.args]]
plot(pred.seq)

writeRaster(pred.seq, "../../pred/out/Mn/Mn_SDM_ME_1km.tif", overwrite = TRUE)
```


### Threshold

```{r Thresh func}
#Function code: https://babichmorrowc.github.io/post/2019-04-12-sdm-threshold/
sdm_threshold <- function(sdm, occs, type = "mtp", binary = FALSE){
  occPredVals <- raster::extract(sdm, occs)
  if(type == "mtp"){
    thresh <- min(na.omit(occPredVals))
  } else if(type == "p10"){
    if(length(occPredVals) < 10){
      p10 <- floor(length(occPredVals) * 0.9)
    } else {
      p10 <- ceiling(length(occPredVals) * 0.9)
    }
    thresh <- rev(sort(occPredVals))[p10]
  }
  sdm_thresh <- sdm
  sdm_thresh[sdm_thresh < thresh] <- NA
  if(binary){
    sdm_thresh[sdm_thresh >= thresh] <- 1
  }
  return(sdm_thresh)
}
```

### Minimum training presence 

```{r MTP}
mtp <- sdm_threshold(pred.seq, occs, type = "mtp")
plot(mtp)

writeRaster(mtp, "../../pred/out/Mn/Mn_SDM_ME_1km_mtp.tif", overwrite = TRUE)
```

### 10th percentile training presence

```{r}
p10 <- sdm_threshold(pred.seq, occs, type = "p10")
plot(p10)

p10.b <- sdm_threshold(pred.seq, occs, type = "p10", binary = TRUE)
plot(p10.b)

writeRaster(p10, "../../pred/out/Mn/Mn_SDM_ME_1km_p10.tif", overwrite = TRUE)
```










